---
title: "Assignment 3 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3

In this assignment we do the following:
- we run a Bayesian meta-analysis of pitch variability in ASD,based on previously published literature
- we analyze pitch variability in ASD in two new studies using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors. 

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Perform a meta-analysis of pitch variability from previous studies of voice in ASD
- N.B. all you need is in the two intro videos
- the data is available as Ass3_MetaAnalysisData.tsv
- You should calculate Effect size (cohen's d) and Standard Error (uncertainty in the Cohen's d) per each study, using escalc() from the metafor package (also check the livecoding intro)
- N.B. for the purpose of the assignment we're only interested in getting a meta-analytic effect size for the meta-analytic prior (and not e.g. all the stuff on publication bias). See a brms tutorial here: https://vuorre.netlify.com/post/2016/09/29/meta-analysis-is-a-special-case-of-bayesian-multilevel-modeling/ The formula is EffectSize | se(StandardError) ~ 1 + (1 | Paper). Don't forget prior definition, model checking, etc.
- N.B. the livecoding video is not perfect, you *can* (but don't have to) improve it: use a t-student likelihood, test the effects of the priors and refine them, check the robustness of results to more or less skeptical priors, etc. 
- Write down the results of the meta-analysis in terms of a prior for step 2.

```{r step 1}
# Importing meta-analysis
pacman::p_load(readr, 
               tidyverse,
               metafor,
               brms, 
               rethinking, 
               viridis, 
               bayesplot)


MA_d <- read_delim("Ass3_MetaAnalysisData.tsv", delim ="\t")
View(MA_d)


MA_d <- MA_d %>% 
  mutate(
    PitchVariabilityASD_Mean = as.numeric(PitchVariabilityASD_Mean),
    PitchVariabilityTD_Mean = as.numeric(PitchVariabilityTD_Mean),
    PitchVariabilityASD_SD = as.numeric(PitchVariabilityASD_SD),
    PitchVariabilityTD_SD = as.numeric(PitchVariabilityTD_SD)
  )
  

# Calculating effect sizes

MA_d <- MA_d %>% subset(!is.na(Paper))
  

MA_d <- escalc(measure = "SMD", #standardized mean difference
               n1i = TD_N,
               n2i = ASD_N,
               m1i = PitchVariabilityTD_Mean,
               m2i = PitchVariabilityASD_Mean,
               sd1i = PitchVariabilityTD_SD,
               sd2i = PitchVariabilityASD_SD,
               data = MA_d,
               slab = Paper)  

colnames(MA_d) # Now also containing yi (effect size) and vi (variance)

MA_d <- MA_d %>% 
  mutate(
    StandardError = sqrt(vi)) %>%
  rename(EffectSize = yi)
      
# Making the formula
MA_f <- bf(EffectSize | se(StandardError) ~ 1 + (1 | Population))  

# Get priors

get_prior(MA_f, data = MA_d, family = gaussian())

MA_prior <- c(
  prior(normal(0,1), class = Intercept),#sceptical prior
  prior(normal(0,.3 ), class = sd) 
)
  
sd(MA_d$EffectSize, na.rm = T)

# Define the model

MA_m0 <- brm(
  MA_f,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = "only",
  chains = 2,
  cores = 2, 
  file = "MA_m0", 
  file_refit = "on_change"
)


pp_check(MA_m0, nsamples = 100)


MA_m1 <- brm(
  MA_f,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = T,
  chains = 2,
  cores = 2,
  file = "MA_m1", 
  file_refit = "on_change"
)

pp_check(MA_m1, nsamples = 100)

summary(MA_m1)

# MA effect mean = -0.43, sd = 0.1
MA_mean <- fixef(MA_m1)[[1]]
MA_SE <- fixef(MA_m1)[[2]]

# Group level effects
# Average error that the population level estimates
MA_heterogeneity = 0.32
```


Step 2: Analyse pitch variability in ASD in two new studies for which you have access to all the trials (not just study level estimates)
- the data is available as Ass3_data.csv. Notice there are 2 studies (language us, and language dk), multiple trials per participant, and a few different ways to measure pitch variability (if in doubt, focus on pitch IQR, interquartile range of the log of fundamental frequency)
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).
- Is there any structure in the dataset that we should account for with random/varying effects? How would you implement that? Or, if you don't know how to do bayesian random/varying effects or don't want to bother, is there anything we would need to simplify in the dataset?

```{r step 2}
#Loading data 
df <- read_csv("Ass3_data.csv")
    
#Cleaning dataframe
df <- df %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(Age = as.factor(Age),
         ID = as.factor(ID)) %>% 
  select(ID, Language, Diagnosis, Gender, Age, starts_with("Pitch"))

#Standardizing all numeric values, centering and scaling
df <- df %>% mutate_if(is.numeric, scale)

#Exploring pitch_IQR, seems like it is distributed as students_t with a long tail 
hist(df$Pitch_IQR)
```

Step 3: Build a regression model predicting Pitch variability from Diagnosis.
- how is the outcome distributed? (likelihood function). NB. given we are standardizing, and the meta-analysis is on that scale, gaussian is not a bad assumption, but check t-student as well. Lognormal would require us to convert the prior to that scale.
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on? Start simple, with Diagnosis only. Add other predictors only if you have the time and energy!
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it, test its predictions (prior predictive checks), its impact on the posteriors (prior posterior updates checks).
- Evaluate model quality. Describe and plot the estimates.

```{r step 3}
#Building model 1
f1 <- bf(Pitch_IQR ~ 1 + Diagnosis + (1|ID))

#Getting priors 
get_prior(f1, data = df)

#Setting priors 
f1_prior <- c(
  prior(normal(0,.5), class = Intercept), #Mean of 0 and sd of .5 (1)
  prior(normal(0,.5), class = sd), #Mean of 0 and sd of .5
  prior(normal(0,0.2), class = b, coef = DiagnosisTD), #Effect size of cohens d .2 (normal effect size)
  prior(normal(.5, .2), class = sigma) 
)

#Model 1 only with sample prior 
m1_prior <- brm(
  f1,
  data = df,
  family = gaussian(),
  prior = f1_prior,
  sample_prior = "only",
  chains = 4, 
  cores = 4, 
  seed = 12, 
  file = "m1_prior", 
  file_refit = "on_change"
)

#Checking 
pp_check(m1_prior, nsamples = 100)

#Better pp_check  
y_pred <- posterior_linpred(m1_prior)
dens(y_pred)

#Model 1 with data 
m1 <- brm(
  f1,
  data = df,
  family = gaussian(),
  prior = f1_prior,
  sample_prior = TRUE,
  chains = 4, 
  cores = 4, 
  seed = 12, 
  file = "m1", 
  file_refit = "on_change"
)

### Model quality check 

#Summary
summary(m1)

#Checking 
pp_check(m1, nsamples = 100)

#Better pp_check 
y_pred1 <- posterior_linpred(m1)
dens(y_pred1)

#Checking the chains 
color_scheme_set("viridis")
mcmc_trace(m1,  pars = "b_Intercept", "b_DiagnosisTD") + theme_classic()
mcmc_rank_overlay(m1, pars = "b_Intercept", "b_DiagnosisTD") + theme_classic() #Both looks fine

#Plot of estimates
mcmc_plot(m1)

#Hypothesis testing
hypothesis(m1, "DiagnosisTD < 0")

#Posterior update check
plot(hypothesis(m1, "DiagnosisTD < 0"))

```
 

Step 4: Now re-run the model with the meta-analytic prior
- Evaluate model quality. Describe and plot the estimates. 
- N.B. you need to assess the meta-analytic informed prior (prior pred checks, prior-posterior update checks) and if relevant you can always change it in motivated ways (e.g. too confident, doesn't let the model actually learn from the data, so increase sd)

```{r}
# MA effect mean = -0.43, sd = 0.1
MA_mean <- fixef(MA_m1)[[1]]
MA_SE <- fixef(MA_m1)[[2]]

# Group level effects
# Average error that the population level estimates
MA_heterogeneity = 0.32

#Setting model 
f2 <- bf(Pitch_IQR ~ 1 + Diagnosis + (1|ID))

#Getting priors
get_prior(f2, data = df)

#Setting the meta analytic priors 
f2_prior <- c(
  prior(normal(0,.5), class = Intercept),
  prior(normal(0,.5), class = sd), 
  prior(normal(-0.43,0.1), class = b, coef = DiagnosisTD), # putting the MA prior here
  prior(normal(.32, 0.1), class = sigma) #Population effects here 
)

#Running model prior only
m2_prior <- brm(
  f2,
  data = df,
  family = gaussian(),
  prior = f2_prior,
  sample_prior = "only",
  chains = 4, 
  cores = 4, 
  seed = 12, 
  file = "m2_prior", 
  file_refit = "on_change"
)

#Checking prior
pp_check(m2_prior, nsamples = 100)

#Better check
y_pred2 <- posterior_linpred(m2_prior)
dens(y_pred2)

#Running model on data
m2 <- brm(
  f2,
  data = df,
  family = gaussian(),
  prior = f2_prior,
  sample_prior = TRUE,
  chains = 4, 
  cores = 4, 
  seed = 12, 
  file = "m2", 
  file_refit = "on_change"
)

### Model quality check
summary(m2)

#Checking prior
pp_check(m2, nsamples = 100)

#Better check
y_pred3 <- posterior_linpred(m2)
dens(y_pred3)

#Checking chains
color_scheme_set("viridis")
mcmc_trace(m2,  pars = "b_Intercept", "b_DiagnosisTD") + theme_classic()
mcmc_rank_overlay(m2, pars = "b_Intercept", "b_DiagnosisTD") + theme_classic()

#Plot of estimates 
mcmc_plot(m2)

#Hypothesis checking
hypothesis(m2, "DiagnosisTD > 0")

#Posterior update check
plot(hypothesis(m2, "DiagnosisTD > 0"))
```

Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare the two models (LOO)
- Discuss how they compare and whether any of them is best.

```{r step 5}
#Comparing with loo
m1 <- add_criterion(m1, "loo", moment_match = TRUE)
m2 <- add_criterion(m2, "loo", moment_match = TRUE)

w_loo <- loo_compare(m1, m2, criterion = "loo")

w_loo
  #Model 2 seems like the best

#Loo weights 
loo_model_weights(m1,m2)
    #Some insecurities about the models 
pp_average(m1, m2)
```


Step 6: Prepare a nice write up of the analysis and answer the questions at the top.
